---
title: "Things to do after installing Ubuntu Linux for Bioinformatics"
author: "Welliton Souza"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

This tutorial is focused on my toolbox for working with genomic data. I manly use R statistical software and Bioconductor packages. It may not fit for everyone.

I tested these commands in Ubuntu Linux version 16.04 LTS 64 bits. Some commands may not work with old versions of Ubuntu.

<kbd>Ctrl</kbd>+<kbd>Alt</kbd>+<kbd>T</kbd> opens the Terminal.

## 1. Update your system

```{bash}
sudo apt-get update
sudo apt-get upgrade -y
```

Reboot your system if necessary.

```{bash}
sudo reboot
```

## 2. Install version control systems

[Git](https://git-scm.com/). [Reference](https://git-scm.com/download/linux).
Git is necessary for working with [GitHub](https://github.com/).

```{bash}
sudo apt-get install git -y
```

[Apache Subversion](https://subversion.apache.org/).
[Reference](https://subversion.apache.org/packages.html#ubuntu)

```{bash}
sudo apt-get install subversion libapache2-svn -y
```

## 3. Install missing command line tools

Install vim.

```{bash}
sudo apt-get install vim -y
```

Install htop.

```{bash}
sudo apt-get install htop -y
```

Install GNU Screen.

```{bash}
sudo apt install screen -y
```

Learn how to use `screen` <a href="screen.html" target="_blank">here</a>.

## 4. Install R and RStudio

First check what is the codename of your Ubuntu installation. [Reference](https://help.ubuntu.com/community/CheckingYourUbuntuVersion).

```{bash}
lsb_release -c
```

[Here](https://wiki.ubuntu.com/DevelopmentCodeNames) is a list of Ubuntu versions and their codenames.

Edit, as superuser, `/etc/apt/sources.list` file adding the lines below at the end of text. There are many [mirrors](http://cran.r-project.org/mirrors.html) available.

```
# R repositories - lasted package versions
deb https://cloud.r-project.org/bin/linux/ubuntu/ xenial/
```

Add a verified key.

```{bash}
sudo apt-key adv --keyserver keyserver.ubuntu.com --recv-keys E084DAB9
```

Update your system package manager. Install R and its system dependencies.

```{bash}
sudo apt-get update
sudo apt-get install r-base-dev -y
```

Install common system packages dependencies for R.

```{bash}
sudo apt-get install libcurl4-openssl-dev ghostscript libxml2-dev libssl-dev -y
```

Install [RStudio desktop](https://www.rstudio.com/products/rstudio/#Desktop) version. Also, there is a [server version](https://www.rstudio.com/products/rstudio/#Server). There are installation files for different Linux distributions and versions, [see more](https://www.rstudio.com/products/rstudio/download3/).

```{bash}
wget https://download1.rstudio.org/rstudio-1.0.44-amd64.deb -O /tmp/rstudio.deb
sudo dpkg -i /tmp/rstudio.deb
```

If installation of RStudio fail, run the command line below. It will install missing system package dependencies.

```{bash}
sudo apt-get -f install -y
```

## 5. Install Bioconductor core packages

Install [Bioconductor](http://bioconductor.org/), on R. This will install the base packages. [Reference](http://bioconductor.org/install/).

```{r}
source("https://bioconductor.org/biocLite.R")
biocLite()
```

Install [tidyverse](https://github.com/tidyverse/tidyverse) package (from RStudio). This will install a bundle of useful package for manipulating general data on R. Below a list of (some) packages that will be installed.

- [ggplot2](http://ggplot2.org/) for data visualisation.
- [dplyr](https://github.com/hadley/dplyr) for data manipulation.
- [tidyr](https://github.com/hadley/tidyr) for data tidying.
- [readr](https://github.com/tidyverse/readr) for data import.
- [purrr](https://github.com/hadley/purrr) for functional programming.
- [tibble](https://github.com/tidyverse/tibble) for tibbles, a modern re-imagining of data frames.

**NOTE:** it is safer using `biocLite` function from `BiocInstaller` package instead of `install.packages` function. 

```{r}
library(BiocInstaller)
biocLite("tidyverse")
```


## 6. Install authoring tools

Install RMarkdown for working with markdown. It is possible to generate HTML, DOCX and PDF (PDF requires LaTeX).

```{r}
biocLite("rmarkdown")
```

Install LaTeX for generating PDF.

```{bash}
sudo apt-get install texlive texlive-latex-extra texlive-fonts-extra texinfo -y
```

## 7. Install package devlopment tools

Install commpilers and system dependencies.

```{bash}
sudo apt-get install build-essential gfortran libreadline6-dev libcurl4-gnutls-dev libpcre3-dev libbz2-dev xorg-dev
```

Install package development tools.

- [devtools](https://github.com/hadley/devtools).
- [roxygen2](https://github.com/klutometis/roxygen).
- [testthat](https://github.com/hadley/testthat).
- [BiocCheck](http://bioconductor.org/packages/release/bioc/html/BiocCheck.html).

```{r}
biocLite(c("devtools", "roxygen2", "testthat", "BiocCheck"))
```

## 8. Install Microsoft Windows fonts

[Reference](https://help.ubuntu.com/community/RestrictedFormats/Microsoft_Fonts).

```{bash}
sudo apt-get install ttf-mscorefonts-installer -y
sudo fc-cache -f -v
```

## 9. Swappiness

By default, this command will return 60.

```{bash}
cat /proc/sys/vm/swappiness
```

Edit, as superuser, `/etc/sysctl.conf` file addin the line below at the end of the text.

```
vm.swappiness = 10
```

Reboot computer.