---
title: "Report template"
author: "Welliton Souza, Elizabeth Borgognoni"
date: "`r Sys.Date()`"
bibliography: "bibliography.bib"
output:
  html_document:
    toc: yes
    toc_float: true
    code_folding: show
    df_print: kable
---

See the [source file](report-template.Rmd) to see the hidden configurations and
code chunks.

# Explaning configurations

- `title` will appear at the top of document;
- `author` the name(s) of author(s);
- `date` keep date of the last compilation;
- `bibliography` path BibTeX file;
- `html_document` generates HTML file;
- `toc` create Table of Contents;
- `toc_float` create froating toc (right side);
- `code_folding` Replace `show` to `hide` to hide R codes and add button to show the code.
- `df_print` use `kable` (from knitr package) to print `data.frame` objects.

# Citing R and Bioconductor package

Add this code chunk before citing a package.

```{r bioc, include=FALSE}
library(BiocStyle)
```

Cite Bioconductor using inline R code, example `r Biocpkg("Rqc")`.
The same for CRAN packages, example: `r CRANpkg("ggplot2")`.
Or packages from GitHub: `r Githubpkg("labbcb/GA4GHclient")`.

# Citing articles

Append BibTex entries to `bibliography.bib` file.
Cite aticles like LaTeX, example: @Wickham2014 or [@Wickham2014].
References will appear ate the end of the document.

# Interactive tables 
```{r}
library(DT)
datatable(mtcars, options = list(scrollX = TRUE))
```

`options = list(scrollX = TRUE)` parameter limits the table by the width of the report area.

# Interactive plots

Select only gear and hp variables.

```{r, message=FALSE}
library(dplyr)
mtcars_tidy <- select(mtcars, gear, hp) %>%
    mutate(gear = as.factor(gear))
```


Plot using `r CRANpkg("ggplot2")` plus `r CRANpkg("plotly")`.

```{r, message=FALSE}
library(ggplot2)
library(plotly)

# Plot using ggplot2 
p <- ggplot(mtcars_tidy, aes(x = gear, y = hp)) +
        geom_boxplot() +
        labs(x = "Number of gears", y = "Horse power") +
        ggtitle("Distribution of horse power by gears")

# Add interactivity using plotly
ggplotly(p)
```

Plot using only plotly package.

```{r}
plot_ly(mtcars_tidy, x = ~gear, y = ~hp, type = "box") %>% 
    layout(title = "Distribution of horse power by gears",
           xaxis = list(title = "Number of gears"), yaxis = list(title = "Horse power" ))
```

# Session info

From devtools package.

```{r sessioninfo-devtools}
devtools::session_info()
```

From base package
```{r sessioninfo-base}
sessionInfo()
```

# References