---
title: "Reproducible report template"
author: "Welliton Souza <well309@gmail.com>"
date: "`r Sys.Date()`"
bibliography: "bibliography.bib"
output:
  html_document:
    toc: yes
    toc_float: true
    code_folding: hide
    df_print: kable
---

# Explaning configurations

- `title` will appear at the top of document;
- `author` the name(s) of author(s);
- `date` keep date of the last compilation;
- `bibliography` path BibTeX file;
- `html_document` generates HTML file;
- `toc` create Table of Contents;
- `toc_float` create froating toc (right side);
- `code_folding` hide R codes and add button to show the code;
- `df_print` use `kable` (from knitr package) to print `data.frame` objects.

# Citing R and Bioconductor package

Add this code chunk before citing package.

```{r bioc, include=FALSE}
library(BiocStyle)
```

Cite Bioconductor using inline R code, example `r Biocpkg("Rqc")`.
The same for CRAN packages, example: `r CRANpkg("ggplot2")`.
Or packages from GitHub: `r Githubpkg("labbcb/GA4GHclient")`.

# Citing articles

Append BibTex entries to `bibliography.bib` file.
Cite aticles like LaTeX, example: @Wickham2014 or [@Wickham2014].
References will appear ate the end of the document.

# Session info

Print session info

From devtools package.

```{r sessioninfo-devtools}
devtools::session_info()
```

From base package
```{r sessioninfo-base}
sessionInfo()
```

# References